Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-09 cordillera-cycle.tex
Work in progress on the discussion.

During the last glacial cycle, geomorphological evidence has allowed to
identify two major advances of the Cordilleran ice sheet. These correspond to
the last glacial maximum \aref, and the older xxx glaciation, found in the
northernmost parts of the ice sheet \aref. It should be noted that all
simulations presented here reproduce these two major events, which constitutes
a first-order agreement between model results and geomorphological evidence.

However, the timing of these events depends highly on the choice of temperature
time-series used to force the model. For instance, the timing of the last
glacial maximum is best reproduced by the Antarctic records, while simulations
driven by the NGRIP and oceanic records pre-date the last glacial maximum by
several thousands of years. Regarding the oceanic records, the early volume
maxima can be explain by an early warming in the palaeo-temperature record.
This early warming has been previously documented, and explained as a
weakening of the California current, precisely associated with the presence of
the Cordilleran ice sheet \aref. This precocious glacial maximum illustrates
the complexity of climate-ice-sheet feedbacks on a regional scale, and shows
that, although closely located to the modelling domain, the ODP~1012 and
ODP~1020 proxy records records can't be reliably used as a temperature forcing
to model the last Cordilleran ice sheet.

>>>>

Independently of the choice of palaeo-temperature record used as forcing
time-series for the ice sheet model, our simulations consistently produce two
glaciation maxima during MIS~4 (61.9--55.4\,\unit{kyr}) and MIS~2
(29.5--16.9\,\unit{kyr}). These events broadly correspond in timing to the
Gladstone (MIS~4) and McConell (MIS~2) glaciation documented by geological
evidence in the northern parts of the model domain, and to the Fraser (MIS~2)
glaciation in the southern part.

However, the timing of the modelled glacial maxima strongly depends on the
choice of palaeo-temperature record used, and a more precise agreement with
timing reconstructed from geological evidence is only attained by some of the
simulations. According to radiocarbon dating, the last glacial maximum advance
of the southern Cordilleran ice sheet margin has been constrained between
00.0\,\unit{kyr} and 00.0\,\unit{kyr}. These dates have been confirmed by the
offshore sediment record, which shows an increase of xxx
betweem 00.0\,\unit{kyr} and 00.0\,\unit{kyr}.

Among simulations presented here, only those forced by the GRIP, EPICA and
Vostok records place the last glacial maximum within this range
(Fig.~\ref{fig:timeseries}, lower panel). Simulations driven by the NGRIP,
ODP~1012 and ODP~1020 records, on the opposite, produce a last glacial maximum
that pre-dates reconstructed ages by several thousand years. Concerning
simulations driven by oceanic (ODP~1012, ODP~1020) records, this early
deglaciation is caused by an early warming in the palaeo-temperature records
(Fig.~\ref{fig:timeseries}, upper panel). However, it has been shown that this
early warming is a local effect, corresponding to a weakening of the California
current \aref. The California currents is a southwards current driving cold
waters along the western coast of USA. It is thought to have weakened in the
presence of the Cordilleran ice sheet, explaining paradoxically warmer
sea-surface temperatures at the location of the ODP~1012 and ODP~1020 sites
during the last glacial maximum \aref.

This precocious glacial maximum illustrates
the complexity of climate-ice-sheet feedbacks on a regional scale, and shows
that, although closely located to the modelling domain, the ODP~1012 and
ODP~1020 proxy records records can't be reliably used as a temperature forcing
to model the last Cordilleran ice sheet.
